\documentclass[a4paper,14pt]{extarticle}

\usepackage{setspace}
\setstretch{1.3}

\usepackage{indentfirst}

\frenchspacing


\usepackage{cmap}

\usepackage{geometry}
\geometry{left=3cm}
\geometry{right=2cm}
\geometry{top=2cm}
\geometry{bottom=2cm}


\begin{document}

\section*{Motivation letter}

!!!
My name is Alexey Grigorev, and as a graduate student I apply to the \textbf{Erasmus Mundus Master's Program in Information Technologies for Business Intelligence (IT4BI)} for the profound study of Business Intelligence and related aspects such as data warehousing, data mining and decision support.

Through both studying and working I gained solid background, ...  and this program exactly matches my study objectives. Having good both academic and professional backgrounds, I am ... And this program ideally suits my interests and study objectives.
This program fits my needs perfectly

Why I? Why my academic background is good? Why professional?

\subsection*{Academic background}

In 2006 I entered Far Easter State Academy for Humanities and Social Studies (afterwards renamed to Priamurian State University) in the city of Birobidzan, Russia, to study computer science and in 2010 I graduated with honors and received bachelor's degree. The courses I took were chiefly devoted to database design for small enterprises: data management, theory of information processes and systems, corporate IT systems and so on. My final graduate thesis was on automated information system for enterprise ``Sterkh'', Birobidzhan. However, disciplines of the most interest for me were mathematical analysis, probability and statistics, and statical analysis of time series.

From the beginning of my studies I realized the importance of the English language and enrolled in an advances training of English and in 2011 graduated with a diploma, equated to bachelor's degree, which allows me to teach the language in general education institutions.

Apart from studies, I was actively engaged in many extracurricular activities. My main research interests during studies were software engineering, resulted in Von Neumann machine model shown in ``Programmer 2009'' conference; image processing, resulted in a paper about copyright protection and watermarking shown in ``Programmer 2010'' conference; and, lastly, applied statistics, with work around cluster analysis and time series procession shown in 57th Scientific Collegiate Conference in Far Eastern State University of Social Studies.

From 2007 to 2010 I annually was a part of PSU's team in Far Eastern Quarter Finals of North-eastern European Regional Collegiate Programming Contest held by ACM. In 2007 and 2008 I participated in individual mathematical contests, where in 2008 showed the best result amongst all PSU participants. Finally, in 2010 I competed in individual nation-wide programing contest where I ranked 4th among participants form Far East region of Russia.

My achievements in studying and extracurricular activities did not go unnoticed and as a recognition in the end of 2009 I received a prestigious governmental award for talented youth. However, among all the tournaments I have partaken in the most memorable was ``the best graduate'' contest in my Alma Mater, between all PSU's graduates from all faculties, and in the finals I faced highly competent and competitive rivals, all with high chances of winning. In the end I managed to win by a narrow margin (and?)

\subsection*{Professional background}

Upon graduation I decided to take a break from studying. There were two reasons for this: firstly, I wanted to acquire more practical knowledge, and secondly, I was driven by desire to find out what particular field is more interesting for me as well as more appealing for business. I moved to Nizhniy Novgorod where I worked in several companies as a software developer.

The first company was developing a system for robotizing e-commerce storages, and in the second one we were in charge of creating a system for selling ringback tones (melodies you hear in the background when you are calling to a mobile phone). The companies are quite different, but there is something that they have in common. Namely, they both need to process large volumes of data efficiently, quickly adapting to the needs of customers, whether it be putting complimentary goods on the same rack or suggesting a tone which a customer is likely to buy.

Now I am working in Poland for an investment back. Trades from many exchanges have to be processed and booked for commission calculation, and the goal of our team is to ensure that the resulting charge is accurate. I _see_ that financial institutions have to deal with large volume of information, every day processing billions and billions of trades. It also proves that data analysis is crucial, and I do my best to keep up with modern trends and technologies in this field.

With technologies developing at a startling pace, being limited only by the current scope of work means falling behind, and I am always looking forward to absorbing new knowledge. For this reason, I attend trainings, take part in local communities (such as Nizhniy Novgorod Java community) or in communities at work (architect and agile community). At the moment I am taking a training about basics of data warehousing and I feel I like it very much. ???


???
Contribution to open source projects is another activity I am proud of???. _As the most significant contribution, _ I played a team lead role in open source project jtalks.org, a project for Russian-speaking Java community -- what it for. Participation in open source has given me much: name, connections, experience in building enterprise-level application from scratch, plus experience in mentoring.

Furthermore, I understand the value of sharing knowledge and keep a professional blog where I post all my thoughts around IT, ?? be it a review of a new technology, summary of a book or a technical article, implementation of algorithms and so on. It started as a place where I could keep notes just for myself but eventually it evolved into a web site which around a hundred of people visit every day.


\subsection*{Study objectives}

With my working experience I have gained a significant insight into the needs of business. From my point of view, in the foreseeable future data scientists - programmers with analytical skills and knowledge of statistics - will be in increasingly high demand. The main reason for this is that over the last decades enterprises have accumulated prodigious amounts of information, and each year the number of data is increased exponentially. Hence, such scientists will be needed desperately to put this data in order.

I am particularly eager to become a data scientists, and in the future I would like to concentrate on the following areas: business intelligence, data warehousing, data mining and big data. While I am constantly learning and self-improving, it is not enough, and I feel a certain lack of skills. For that reason, in order to be successful in these areas, a __good and intensive education__ is needed, and such an __outstanding__ master program will definitely help me to reach my goals.

Therefore, I am deeply interested in the courses \textbf{IT4BI} provides, because the curriculum fits perfectly my educational objectives. I am especially willing to develop decision support and data mining skills, and that is why I would like to choose Decision Support and Business Intelligence as a specialization. !!! Google about professors.

Additionally, I am highly interested in cultural exchange: traditions, art, and local cuisine really appeal to me. With this program, I will have an opportunity to live in Belgium and France, visit Rene Magritte museum in Bruselle? and Luvr in Paris. But... is architecture. I will definitely visit Bruges in Belgium and will take a trip to see the castles of the Coire Valley in France. Moreover, it will give me such a great opportunity to learn French. I like how the French language sounds, but unfortunately I have never had an opportunity to learn the language and know nothing about it. I hope with this program I will be able to reach at least lower-intermediate level. ... I am looking forward to visiting both countries.

???
Upon finishing the program I plan to pursue PhD and combine professional and teaching activities. The latter is especially important for me, because besides building my career I am extremely interested in sharing my experience and knowledge with others. After graduating I will definitely have a lot to share... The program definitely will give me a solid??? background??? for this.

???
Linker? The reasons why I am a good fit for the scholarship are good experience, both academic and professional, crave for constant improving, proactiveness and open-mindedness. With my strong academic background, relevant working experience and constant crave for knowledge, I am confident that I am qualified and able to perform well in this program.

Thank you very much for considering my application. I am looking forward to your positive response.

\end{document}
