\documentclass[a4paper, 12pt]{article}

\usepackage{indentfirst}
\frenchspacing

\usepackage{cmap}

\usepackage{geometry}
\geometry{left=1.7cm}
\geometry{right=1.7cm}
\geometry{top=1.5cm}
\geometry{bottom=1.7cm}

\newcommand{\group}[1] {\subsection{#1}} 
\newcommand{\discipline}[1] {\textbf{#1} \\ }
\newcommand{\desc}[1] { #1 \\ }

\begin{document}

\group{General humanitarian and socio-economic disciplines}


\discipline{Foreign language (English)}
\desc{sufficient to achieve the B2 level of a foreign language, including all of the skills: speaking, reading, writing, listening and grammar. The Culture of the countries where the language is spoken.}

\discipline{Physical culture}
\desc{Elements of the healthy life, and how to optimize personal performance. General physical training. Sports.}

\discipline{History of Russia}
\desc{History as a science, Russian history - as an integral part of world history. History of Russia from the ancient times to the present. Culture.}

\discipline{Political science}
\desc{Functions of Political Science. The role and the place of politics in modern societies. The function of social laws and regulations.}

\discipline{Sociology}
\desc{Sociology as a science, its functions. Social action and social interaction, culture, social communication, social groups and structures, family.}

\discipline{A man in a cultural context (CPV)}
\desc{What is the culture, its history, typology, culture and identity. What is the place of men in the culture.}

\discipline{Jurisprudence}
\desc{The state and the law, their role in society, main legal systems. International law as a particular system of law.}

\discipline{Russian Language and Speech}
\desc{Styles of the modern Russian written language. Language norms, they role in the formation and functioning of the language. Key issues in improving the literacy skills of speaking and writing.}

\discipline{Philosophy}
\desc{The subject of philosophy. The place and role of philosophy in culture. The history of philosophy, guidelines and major philosophical schools.}

\discipline{Economics}
\desc{The introduction to economics. Goods, needs and resources. Economic systems. Microeconomics.}

\discipline{Safety of life}
\desc{People and the environment, criteria for comfort. The dangers of technical systems. Means of reducing the harmful effects of traumatic and technical systems. Emergency situations.}


\group{Natural scientific component}

\discipline{Physics !!!}
\desc{Physical basis of mechanics. Waves, molecular physics and thermodynamics, electricity and magnetism, optics, atomic and nuclear physics, physics practicum.}

\discipline{Chemistry}
\desc{Chemical systems, chemical thermodynamics and kinetics, reactivity of substances, chemical identification; workshops.}

\discipline{Ecology}
\desc{The biosphere and the man. Global environmental problems, environmental principles of rational use of natural resources and protection of the environment.}

\group{Mathematical component}

\discipline{Algebra and Geometry !!!!}
\desc{Algebra: basic algebraic structures, vector spaces and linear transformations.
Geometry: analytic geometry, multi-dimensional Euclidean geometry}


\discipline{Discrete Mathematics}
\desc{Logical calculations, graphs, theory of algorithms, combinatorics.}

\discipline{Algorithms for Discrete Mathematics}
\desc{Algorithm, analysis of algorithms, the Turing machine, combinatorics, graphs, depth-first and breadth-first searches.}

\discipline{Logic}
\desc{propositional logic, logical consequence, the principle of deduction, predicate logic, syntax and semantics of predicate logic, the principles of logic programming}

\discipline{Computational Mathematics}
\desc{numerical solution of mathematical problems, interpolation and approximation, solving systems of linear and nonlinear equations, differentiation and integration of functions, Fourier transform, error estimation.}

\discipline{Mathematical analysis}
\desc{differential and integral calculus, elements of the theory of functions and functional analysis, differential equations.}

\discipline{Transformation of Fourier and Laplass}
\desc{The basic concepts of complex analysis, Laurent series. zeros of an analytic function. residues and their applications Fourier series. Fourier integral, Laplace transform.}

\discipline{Probability and Statistics !!!}
\desc{Mathematical foundations of probability theory, the model of random processes, hypothesis testing, statistical methods of data processing.}

\discipline{Statistical analysis of time series}
\desc{Linear models. The method of least squares. Time series with discrete and continuous time. Computer simulation and analysis of time series. Programming and modeling in MATLAB and FORTRAN}

\group{By specialty}

\discipline{Informatics}
\desc{Informatics and information processes, the basics of programming. Laboratory workshop in Pascal.}

\discipline{Electrical and electronics}
\desc{Electric and magnetic circuits. Basic definitions, calculation methods of electrical circuits. Analysis and calculation of the chains. Electromagnetic devices. Basic electronics and electrical measurements. Element basis of modern electronic devices.}

\discipline{Metrology, standardization and certification}
\desc{Theoretical basis of metrology. The notion of measurement assurance. Organizational, scientific and methodological basis of measurement assurance. Precision parts, components and mechanisms, certification, and its role in improving the quality of products.}

\discipline{Information Technology !!!!!!!}
\desc{General classification of information technologies and models of processes of transfer, processing and storage of data in information systems, types and characteristics of information technology, the means of implementation.}

\discipline{The theory of information processes and systems !!!}
\desc{The main goal of systems theory, the concept of an information system; system analysis, species descriptions of information systems and models of information systems, the synthesis and decomposition of information systems, information models for decision-making;}

\discipline{Data management}
\desc{The basic concepts of databases and knowledge base management system (DBMS), database architecture, database design, representation of data structures in computer memory, review of commercial DBMS;}

\discipline{Information networks !!!}
\desc{Basic concepts of information networks, models and structure of information networks, methods switching information, security information, network software and hardware networking.}

\discipline{Fundamentals of Control Theory}
\desc{Management and Information Technology, general principles: stability, controllability and observability, mathematical models of objects and systems and controls and presentation models, methods of analysis and synthesis of control systems. Digital control systems, the software implementation of control algorithms in digital systems.}

\discipline{System Modeling}
\desc{The basic concepts of the modeling theory, types of modeling; simulation models of information processes, mathematical modeling. formalization and algorithmic information processes, conceptual models, simulation modeling of information systems and networks. Modeling in GPSS World.}

\discipline{Computer Architecture and Systems !!!}
\desc{Von Neumann architecture, CPU and RAM; interruptions, I/O; peripherals, architectural features, parallel systems, information and computer systems and networks.}

\discipline{Operating Systems}
\desc{Design principles of operating systems, the basic functions, a review of modern operating systems, the management of the computing process, I/O, real memory, virtual memory; how to build an OS}

\discipline{Programming Technology !!!}
\desc{The main stages of problem solving on a computer; quality criteria of programs; program life cycle, the problem and the specification of the program, and the program in high-level, standard data types.}

\discipline{Geometry and computer graphics}
\desc{Computer graphics, geometric modeling, graphics primitives and attributes, a presentation of video and machine generation, basic graphics;}

\discipline{Representation of knowledge in information systems !!!}
\desc{A logical model of knowledge representation and inference rules, the theoretical fundamentals, theory and technique of knowledge acquisition, expert system. languages ​​of artificial intelligence, logic programming model, practice in Lisp and Prolog.}

\discipline{Information security and data protection !!!}
\desc{The purpose, access control systems, methodology, telecommunications and network security, cryptography basics, watermarking as a means of information protection. Laboratory practice, cryptography C.}

\discipline{Corporate Information Systems !!! Enterprise!}
\desc{The structure of corporations and enterprises, the architecture of enterprise information systems (EIS), the application of EIS, modeling and designing, CIS programming, examples of KCS.}

\discipline{Administration in Information Systems}
\desc{Functions, procedures and administration services, configuration management services, accounting and security, operation and maintenance of information systems, programming systems administration;}

\discipline{Intelligent information systems}
\desc{The basic concepts of artificial intelligence, information and data, expert systems. Information models of knowledge, semantic network, data processing rules. The methods of knowledge representation in databases of information systems.}

\discipline{Multimedia Technology !!!}
\desc{The concept of multimedia technology, classification and application of multimedia applications, types and file formats: text, images, audio, video. Virtual reality. Software, technology for creating multimedia products;}

\discipline{Reliability of information systems !!!}
\desc{The basic definitions of reliability, fault classification, characteristics of reliability, reliability testing, and methods of increasing the reliability of information systems, the impact of the human operator for the operation of information systems.}

\discipline{The reliability of automated systems !!!}
\desc{Basic concepts and components of reliability. Reliability CAM, Reliability restorable and renewable systems. Reliability and quality of the software}

\discipline{Information systems !!!}
\desc{General characteristics of the IC design process, functional model of a user interface, the structure of software modules, development of algorithms, performance analysis, project management: project documentation, design tools, IP, typing design solutions; Information system in Delphi.}

\discipline{High-level programming techniques}
\desc{The basic concepts of programming. The basic methodology of software engineering. Programming languages ​​and systems. Main paradigms: Modular programming, Object-oriented programming. Debugging and testing programs.}

\discipline{Automata Theory !!!}
\desc{The concept of the state machine. The basic definitions. Of automatic functions. Deterministic function. Types of machines: Finite Automata, Structural machines, linear automata.}

\discipline{Mathematical Information Technology !!!}
\desc{An overview of computer mathematics: classification and structure, user interface, data types of mathematical systems, syntax: operators, functions, and expressions. On the example of MathCAD, Maple, Matlab}


\group{Others}

\discipline{Freshman adaptation system}
\desc{}

\discipline{Engineering Psychology !!!}
\desc{As the result of an error, the theoretical analysis of professional activities. The concept of professional tasks simple and complex tasks in the work of the operator. Cognitive process in the operator's problem solving. Command and joint activities.}

\discipline{Technology careers !!!}
\desc{Career, the basic concepts. The study of the labor market. Means of obtaining information about the labor market and how vocational education. Self-presentation. Prepare a resume.}

\discipline{Fundamentals of Scientific and Engineering Research !!!}
\desc{based on research and engineering creativity. Search, storage, and processing of scientific and technical information. Experiment. Processing of the results of experimental studies. Statistical analysis of the primary experimental data.}

\end{document}
