Data Analysis Assigment 1: Loans Data Analysis

========================================================

Preliminaries
-------------------------

```{r loadData}
default.par = par()
setwd("C:/Users/Alexey Grigorev/Documents/courses/coursera/Data Analysis/week3/analysis/my")

loans = read.csv("data/loansData.csv")

summary(loans)

sapply(loans[1,], class)
# note that Interest.Rate abd Debt.To.Income.Ration are characters
# so let's transform them into numeric

loans$Interest.Rate = as.numeric(sub(pattern="%", replacement="", 
                                     loans$Interest.Rate)) / 100

loans$Debt.To.Income.Ratio = as.numeric(sub(pattern="%", replacement="", 
                                     loans$Debt.To.Income.Ratio)) / 100


# also there are some NAs, lets remove them
loans = loans[complete.cases(loans), ]

# 2 observations with NA gone
```

Exploratory analysis
---------------------


```{r}
attach(loans)

hist(Interest.Rate, breaks=50)
quantile(Interest.Rate)
```

### Amount.Requested

```{r}
hist(Amount.Requested)
quantile(Amount.Requested)
```

let's play and see what may affect Interest.Rate
```{r}
plot(Interest.Rate ~ Amount.Requested)
```
no visible pattern

```{r}
plot(Interest.Rate ~ Amount.Requested, col=rgb(0, 0, 1, 0.1), pch=19)
smoothScatter(Interest.Rate ~ Amount.Requested)
```
there seems to be several densest spots.


```{r}
cor(Interest.Rate, Amount.Requested)
```
and correlation is not very high

let's try to cut requested amount into 5 intervals
```{r}
Amount.Requested.Cut = cut(Amount.Requested, breaks=5)
# giving meaningful names to levels
levels(Amount.Requested.Cut) = 
  as.character(tapply(Amount.Requested, Amount.Requested.Cut, median))
    
boxplot(Interest.Rate ~ Amount.Requested.Cut, varwidth=TRUE)
```
Now it's apparent that the bigger requested amount, the higher is the interest rate

### Loan Length

```{r}
boxplot(Interest.Rate ~ Loan.Length, varwidth=TRUE)
```
for longet loans interest rate tend to be higher. apparent relation



### Loan Purpose
```{r}
par(mar=c(9, 3, 1, 1))
boxplot(Interest.Rate ~ Loan.Purpose, pars=list(las=2), varwidth=TRUE)
```
there seems to be a pattern. let's try to sort by medians

```{r}
sort.by.medians = function(what, fac) {
  # from computing for data analysis, week 3 assigment 3

  medians = tapply(what, fac, median)
  # order level names of the factor
  ordered.levels = levels(fac)[order(medians)]
  # and order the factor by ordered names
  ordered.by.medians = ordered(fac, ordered.levels)
  ordered.by.medians
}

boxplot(Interest.Rate ~ sort.by.medians(Interest.Rate, Loan.Purpose), 
        pars=list(las=2), varwidth=TRUE)
par(mar=default.par$mar)
```
so renewable evergy, education and car are with the lowest interest rate; dept consolidation is with the highest

### State

```{r}
boxplot(Interest.Rate ~ State)
boxplot(Interest.Rate ~ sort.by.medians(Interest.Rate, State),
        pars=list(las=2))
```
also in some states median of interest rate is lower than in others

### Debt-To-Income Ratio
```{r}
plot(Interest.Rate ~ Debt.To.Income.Ratio)
smoothScatter(Interest.Rate ~ Debt.To.Income.Ratio)
cor(Interest.Rate, Debt.To.Income.Ratio)
```
no visible pattern and correlation is very low

```{r}
hist(Debt.To.Income.Ratio)
```
seems like a normal distribution

lets try cutting it and see the interest rate
```{r}
Debt.To.Income.Ratio.Cut = cut(Debt.To.Income.Ratio, breaks=9)
levels(Debt.To.Income.Ratio.Cut) = 
  as.character(tapply(Debt.To.Income.Ratio, Debt.To.Income.Ratio.Cut, median))

boxplot(Interest.Rate ~ Debt.To.Income.Ratio.Cut, 
        pars=list(las=2), varwidth=TRUE)
```
now even without ordering it's obvious that the higher the debt/income ration the higher interest rate


### Home Ownership
```{r}
boxplot(Interest.Rate ~ Home.Ownership)
```
doesn't seem to be relevant at all


### Monthly Income
```{r}
plot(Interest.Rate ~ Monthly.Income)
plot(Interest.Rate ~ log10(1 + Monthly.Income), col=rgb(0, 0, 1, 0.2), pch=19)
```
no visible pattern

```{r}
smoothScatter(Interest.Rate ~ log10(1 + Monthly.Income))
```
it seems that it does depend in many cases, but the veriance is too high

### Monthly Income
```{r}
cor(Interest.Rate, Monthly.Income)
```
no correlation

```{r}
boxplot(Interest.Rate ~ cut(log10(1 + Monthly.Income), breaks=15))
```
there seems to be no pattern


### Open CREDIT Lines
```{r}
boxplot(Interest.Rate ~ Open.CREDIT.Lines)
```
for small values it seems to make no difference, however, for biffer IR does get bigger

```{r}
hist(Open.CREDIT.Lines)

Open.CREDIT.Lines.Cut = cut(Open.CREDIT.Lines, breaks=15)
boxplot(Interest.Rate ~ Open.CREDIT.Lines.Cut)

Open.CREDIT.Lines.CutL = cut(Open.CREDIT.Lines, breaks=7)
boxplot(Interest.Rate ~ Open.CREDIT.Lines.CutL, varwidth=TRUE)
```
so it indeed does (we can omit last 2 values as they are not frequent)


### Revolving.CREDIT.Balance
```{r}
smoothScatter(Interest.Rate ~ log10(1 + Revolving.CREDIT.Balance))
cor(Interest.Rate, Revolving.CREDIT.Balance)
```
no correlation, no visible pattern

```{r}
boxplot(Interest.Rate ~ cut(Revolving.CREDIT.Balance, breaks=10))
boxplot(Interest.Rate ~ cut(log10(1 + Revolving.CREDIT.Balance), breaks=10))
```
doesn't seem to have a pattern


### Inquiries in the Last 6 Months

```{r}
table(Inquiries.in.the.Last.6.Months)

boxplot(Interest.Rate ~ Inquiries.in.the.Last.6.Months, varwidth=TRUE)
```
probably there is a trend:
$$0 < 1 < 2 < 3$$. But why 4, 5 and 7 are lower? (seems the reason is that they are rare)

### Employment Length
```{r}
boxplot(Interest.Rate ~ Employment.Length)
```
# seems no pattern


### FICO range
```{r}
boxplot(Interest.Rate ~ FICO.Range)
```
obviously there is a pattern: the higher FICO score is, the less interest rate is

```{r}
detach(loans)
```

so the following variables seem to have influence on Interest.Rate:
* Amount.Requested
* Loan.Length
* Loan.Purpose
* State (seems needn't take into accont)
* Debt.To.Income.Ratio
* Open.CREDIT.Lines (trend is weak)

Modelling
------------

let's concentrate on variables that seem to matter
```{r}
loans1 = loans[, c("Interest.Rate", "Amount.Requested", "Loan.Length", 
          "Loan.Purpose", "Debt.To.Income.Ratio")]

attach(loans1)

lm1 = lm(Interest.Rate ~ Amount.Requested)
plot(Interest.Rate ~ jitter(Amount.Requested), col="blue", pch=19, cex=0.3)
abline(lm1)

plot(Interest.Rate, lm1$residuals, col = Loan.Length, pch = 19, cex=0.7)
```

appears to be non-random pattern here

```{r}
plot(Interest.Rate, lm1$residuals, col = Loan.Purpose, pch = 19, cex=0.7)
```

harder to say if there is a pattern


Now fit a model with factor adjustment for loan purpuse, loan length, and debt-to-income raion

```{r}
lm2 = lm(Interest.Rate ~ Amount.Requested + Loan.Length + 
           Loan.Purpose + Debt.To.Income.Ratio)

plot(Interest.Rate, lm2$residuals, col = Loan.Length, pch = 19, cex=0.7)
```
 far better!

```{r}
plot(Interest.Rate, lm2$residuals, col = Loan.Purpose, pch = 19, cex=0.5)
plot(Interest.Rate, lm2$residuals, col = Debt.To.Income.Ratio.Cut, pch = 19, 
     cex=0.5)
```
also some patterns

estimates:
```{r}
summary(lm2)
```

correlation:
```{r}
sqrt(summary(lm2)$r.square)
```

confidence intervals
```{r}
confint(lm2)
```

Comparison with FICO

```{r}
FICO.Range.min = as.numeric(sapply(strsplit(as.character(loans$FICO.Range), "-"),
                                   function(x) x[1]))

lm.fico = lm(Interest.Rate ~ FICO.Range.min)
summary(lm.fico)
sqrt(summary(lm.fico)$r.squared)

plot(Interest.Rate ~ jitter(Amount.Requested), col = "blue", pch = 19, cex = 0.3)
abline(lm2)
abline(Interest.Rate, lm.fico$fitted.values)
```



Figures
-------

```{r}

library(RColorBrewer)
## Set up a function that makes colors prettier
mypar <- function(a = 1, b = 1, brewer.n = 8, brewer.name = "Dark2", ...) {
    par(mar = c(2.5, 2.5, 1.6, 1.1), mgp = c(1.5, 0.5, 0))
    par(mfrow = c(a, b), ...)
    palette(brewer.pal(brewer.n, brewer.name))
}

## Set size of axes
cx = 1.3

```


```{r fig.width=12, fig.height=4}
mypar(mfrow = c(1, 3))

boxplot(Interest.Rate ~ Amount.Requested.Cut, varwidth = TRUE,
        main="(a)",
        xlab="amounts requested",
        ylab="interest rate",
        col=brewer.pal(5, name="Dark2"))

plot(Interest.Rate, lm1$residuals, col = Loan.Length, pch=19,
     main="(b)",
     xlab="no adjustment resuduals",
     ylab="interest rate")

plot(Interest.Rate, lm2$residuals, col = Loan.Length, pch=19,
     main="(c)",
     xlab="full model resuduals",
     ylab="interest rate")

mypar(mfrow = c(1, 1))
```

