\documentclass[a4paper,12pt]{article}

\usepackage{indentfirst}
\frenchspacing

\usepackage{cmap}

\usepackage{geometry}
\geometry{left=1.7cm}
\geometry{right=1.7cm}
\geometry{top=1.5cm}
\geometry{bottom=1.7cm}

\begin{document}

\section*{Letter of Intent}

My name is Alexey Grigorev, and as a graduate I apply to the \textbf{Master's program in Machine Learning} in the \textbf{KHT Royal Institute of Technology} for the fundamental study of Machine Learning and related aspects such as Artificial Intelligence and Time Series Analysis. I am fully qualified for the studies, and with my strong academic and professional background I am a good match for the program. 

In 2006 I entered Far Eastern State Academy for Humanities and Social Studies (now Priamurian State University) in the city of Birobidzan, Russia, to study computer science and in 2010 I graduated with honors and received bachelor's degree. The courses I took were chiefly devoted to database design for enterprises: Data Management, Theory of Information Processes and Systems, Corporate IT Systems and so on. However, the disciplines I was mostly interested in were Mathematical Analysis, Probability and Statistics and Statical Analysis of Time Series.

Upon graduation I moved to Nizhniy Novgorod where I worked for several companies as a software developer. The first company was developing a system for robotizing e-commerce warehouses, and in the second one we were responsible for creating a system for selling ringback tones. Those companies were quite different, but there was something that they have in common. Namely, they both needed to process large volumes of data efficiently, quickly adapting to the needs of customers, whether it be putting complimentary goods on the same rack or suggesting a tone which a customer is likely to buy.

Now I am working in Poland for an investment bank. Trades from many markets have to be processed and booked for commission calculation, and the goal of our team is to ensure that the resulting charge is accurate. I am able to see that financial institutions have to deal with large volume of information, every day processing billions and billions of trades.

Therefore, trough my working experience I have gained a significant insight into the needs of business. From my point of view, in the not too distant future programmers with machine learning and data analysis skills will be in increasingly high demand. The main reason for this is that over the last decades enterprises have accumulated prodigious amounts of information, and each year the number of data doubles. Hence, such professionals will be needed desperately to enable business to glean important information from this data.

I am particularly eager to gain this kind of expertise, and in the future I would like to focus on the following areas: machine learning, artificial intelligence, data mining and advanced statistics. While I am constantly learning and self-improving, it is not enough, and I feel a certain lack of skills. For that reason, in order to be successful in these fields, a good university education is needed, and such a remarkable master program will definitely help me reach my goals.

Therefore, I am highly motivated to take the courses this program provides, because the curriculum fits my educational objectives perfectly. I am especially willing to develop data analysis skills, and that is why I would like to choose \textbf{Information Retrieval} specialization.

Additionally, I am deeply interested in cultural exchange: traditions, art, and local cuisine really appeal to me, and with this program, I will be able to immerse myself in the Swedish culture, and throughout studying I hope to reach at least lower intermediate level of proficiency in the language.

To summarize, with my strong academic background and relevant working experience, I am confident that I am qualified and able to perform well in this program.

Thank you very much for considering my application. I am looking forward to your positive response.

\end{document}