\documentclass[a4paper,12pt]{article}

\usepackage{indentfirst}
\frenchspacing

\usepackage{cmap}

\usepackage{geometry}
\geometry{left=2cm}
\geometry{right=2cm}
\geometry{top=2cm}
\geometry{bottom=2cm}

% hyperlinks
\usepackage{hyperref}

\begin{document}

\section*{Motivation letter}

Intro?

My name is Alexey Grigorev, and as a graduate I apply to the \textbf{Erasmus Mundus Master's Program in Data Mining and Knowledge Management (DMKM)} for the fundamental study of Data Mining and related aspects such as Data Analysis, Machine Learning and Statistical Modeling. I am fully qualified for the program, and with my strong academic and professional background I am a good match for the scholarship. Not only is the curriculum attractive for me, but the general idea of the Erasmus Mundus program of cultural exchange trough mobility and academic cooperation is also appealing.

\subsection*{Summary}

``Stock exchange mood analyzer'' is an artificial intelligence application for predicting possible price changes on stock exchange markets by doing extensive analysis of the content of financial newspapers, magazines as well as blog posts, twitter and facebook activity. The predictions are afterwards compared against real data, thus leading to a constant cycle of learning on mistakes and improving accuracy of hypotheses.

The product is a web-portal (SaaS) with possibility to monitor the changes of given stack shares and informing the mood of each.

The ultimate goal of the application is to provide real-time, reliable and fault-tolerant, scalable software as a service with access from everywhere.
Model: SaaS (Software as a Service) - web portal.
highly distributed, fault-tolerant

Target-customers are business-oriented people, ... stock exchange players, investors, traders.

The system of that kind is apparently in a high demand. Traditional statistical analysis software have failed to predict crises etc and essentially it is mood that drives the market forward - and by analyzing it and react appropriately it ...

Possible future services:

to scan potentially appealing businesses which are growing in a given field and suggest it as for investment. For example, if someone is looking where to invest, and there is ongoing buzz? about company X - then the company X is probably is good for investing money.

The system may be extended more and to be taught to look for rises and falls in currencies, depending on what a county does. (a conflict may mean a decrease, for example).


Study at ... will help me acquire all needed skills, build needed connections and ...

\newpage

\subsection*{Details}

Many traders and investors rely on newspapers, sometimes even on significant people in twitter, facebook or google plus. Based on opinions from these sources, they decide whether to invest in a certain stock(share) or not. Therefore, in general, ``bad'' mood about a company is likely to lead to price decline as more people will be interested to sell their shares, and vise versa, if a company is praised, ``good'' mood will cause increases as more and more people are interested to invest in the company.

However, this activity can be automated and a machine can be taught to analyze the mood found in that these sources, and to discover whether it is positive or negative and constantly learn and test hypothesis against real data from stock exchanges - find patters between the mood and real trading data, and correcting wrong hypothesis and statistics of wrong guesses. The more data processed, the more accurate predictions are made, and over time it is expected to react appropriately in real time (immediately).

Hence, by scanning these opinions and looking for patters by comparing hypothesis(plural?) against real data from stock exchanges afterwards to correct them, a computer system can be taught. And the more data processed, the more accurate predictions are expected. 

For example, when UBS recently announced job cuts\footnote{\url{http://www.reuters.com/article/2012/10/30/us-ubs-restructure-idUSBRE89S0DM20121030}}, it attracted investors \footnote{\url{http://www.bloomberg.com/news/2012-10-28/ubs-said-to-plan-10-000-job-cuts-investment-bank-shrinks.html}} \footnote{\url{http://wbponline.com/Articles/View/9545/ubs-shares-rise-on-10k-job-cuts-in-restructuring}}, and as result, the price has been growing since October 2012\footnote{\url{http://goo.gl/RSsCE}}.

With software like this all information can be processed automatically, with no human interference at all.

However, handling .... 

key: artificial intelligence, machine learning, distributed calculation, fault-tolerant, add some more buzzwords here


As the goal is to be easily accessible to the end user, it has to be SaaS-system, working on a cloud, 

The amounts of data to process are prodigious, and therefore a distributed system for handling big data is needed. 

\subsection*{Market analysis}

Stock exchanges prices analyzers have existed for a long while ? But the general idea in that kind of software is statistical analysis of past trading activity. Although it is certainly applicable in general, it has a flaw and cannot predict the behavior in crisis or react on certain strategies conducted by conpanies' management. Therefore it has be taken into account and usually people do this by reading. 

When querying google for ``Stock price predictor'', it returns something, but overall the result comprises of articles on how to predict and links, not software.

However there are some programs \footnote{\url{http://dmoz.org/Business/Investing/Research\_and\_Analysis/Software}}. Among them are such programs as GMDH shell \footnote{\url{http://www.gmdhshell.com/}}, IIQ? (Check) System or Trading Expert Pro. While it is certainly a good piece of software, they make predictions based on only the past data from exchanges using old (albeit proven) statistical algorithms. And on top of that, it is a desktop application, not a web-based cloud application.

most of them do analysis bases on only data from se. it's good, but time series analysis and things like this are not good for predicting crisis, people behavior
Found patterns and past data may not reflect current ongoing affairs (discussed in press) and therefore not always good predictors. So this system may be seen as an addition to the aforementioned systems, not as a competitor.

Market is currently growing and the system has to be scalable

(To target)
Current opportunities are used by business intensively. For example, in UBS, Deutche Bank or any other investment financial institution traders rely on ... and likely yo pay or even invest into this program.


\subsection*{Marketing plan}

Once the system is developed, (1.0 version with the most significant functions), it can be presented to major players

and to individuals.
At first, for latter it can be free subscription with possibility to watch only limited number of companies (shares?) and with paid subscription with no limitations.
Ad on specialized forums, google ad, etc.
Quality: high, since it is cloud-oriented, SaaS, distributed - clients can rely on it.

Kay factors on prices would be prices offered by companies with similar offers (and the speed customers want to get updates). - as the system may need tuning for adjusting to an issue of a specific newspaper (i.e. financial times etc)

But the major source of income should be key players who are expected to invest in the software - in order to tailor it to their needs. (and eventually may even acquire the company)

\subsection*{Orientation plan (? Vision?)}

First and foremost, 0.1 version has to be released. Depending on the number of people involved and their skills it could take up to 2 years (for one person without all needed knowledge, skills and experience). At this stage, the project requires extensive research, collecting data, writing a master dissertation etc.
acquiring needed knowledge, skills, making connections, collecting data, making intensive research around it.

(What the Uni can give me goes here!!!)

EIT ... master school provide invaluable background for this
First results are expected, narrowing? broadening the target auditory.

The next year should be aimed on (at) collecting data, calibrating the system to making good results, looking for investors. As a result at this stage, version 1.0 is expected as well and some customers and investors. Tailoring the system to the needs of the business.

Finally, in five years it should be a well-established product with its
% (надежный)
customers etc and key players interested in the product.

\subsection*{Study objectives}

Trough my working experience I have gained a significant insight into the needs of business. From my point of view, in the not too distant future data scientists -- programmers with data mining skills and good knowledge of statistics -- will be in increasingly high demand. The main reason for this is that over the last decades enterprises have accumulated prodigious amounts of information, and each year the number of data doubles. Hence, such professionals will be needed desperately to put this data in order.

I am particularly eager to become a data scientists, and in the future I would like to focus on the following areas: data mining, data warehousing and advanced statistics. While I am constantly learning and self-improving, it is not enough, and I feel a certain lack of skills. For that reason, in order to be successful in these fields, a good university education is needed, and such a remarkable master program will definitely help me reach my goals.

Therefore, I am highly motivated to take the courses \textbf{DMKM} provides, because the curriculum fits my educational objectives perfectly. I am especially willing to develop modeling skills and strengthen further data mining skills, and that is why I would like to choose \textbf{Data Mining and Complex System Modeling} specialization for the second semester and \textbf{Statistical Modeling and Data Mining} specialization for the third.

Additionally, I am deeply interested in cultural exchange: traditions, art, and local cuisine really appeal to me, and with this program, I will be able to immerse myself in French and Spanish cultures. Architecture is my favorite form of art, and I will definitely explore the Loire Valley in France and see the Sagrada Familia church in Barcelona. Moreover, it will give me an exciting opportunity to learn French and Spanish. Both languages sound equally attractive for me, and I hope to reach at least lower intermediate level of proficiency in both.

Upon finishing the program I plan to pursue PhD and combine professional and teaching activities. The qualification obtained through the studies will allow me to easily enroll in any doctoral program, and I am specifically interested in researches conducted by Universit\'e Libre de Bruxelles and Pozna\'n University of Technology. Teaching is also important for me, because besides building my career I am enthusiastic about sharing my experience and knowledge with others.

To summarize, with my strong academic background, relevant working experience and a constant desire for acquiring knowledge, I am confident that I am qualified and able to perform well in this program.

Thank you very much for considering my application. I am looking forward to your positive response.


\end{document} 